CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-04 codecheck-paper.tex
about to read

\textbf{Who's got time for more peer review?} Agree.
The time one spends performing CODECHECKs will likely reduce the time an individual has to work as a reviewer.
However, a CODECHECK is different from peer review.
First, the \textbf{technical nature} of a CODECHECK can provide a more 
clearly defined scope and expectations and thereby time budget compared to conventional 
peer review. 
For example, authors are told what to provide and the codechecker can be told when to stop.
The option for the codechecker to ask for clarification or more 
material can greatly reduce the digging required and thereby
increase \textbf{efficiency}.
Second, the specific skill set of a codechecker allows for \textbf{different groups} to take part in the review process.
ECRs might be attracted to learn more about recent methods, peer review, and reproducibility practices.
Research software experts (Research Software Engineers, RSE\footnote{https://en.wikipedia.org/wiki/Research_software_engineering}) who might
not regularly be involved in writing or reviewing papers might be interested in increasing their connection with scholarly practices.
An extra codechecker may simplify the matchmaking an editor has to do
when identifying suitable reviewers for a submission, as technical and
topical expertise can be provided by different people who do not even have to be traditional peer reviewers.
In fact, the codechecker could even be a software professional employed by a publisher.
Third, recall that we expect CODECHECKs not to be anonymous and always
publicly deposited, which is common practice for reviews of few journals only.
With code and workflows, there also is a good chance that a codechecker's
feedback can directly impact and improve the author's 
work, which may make for a rewarding experience or even provide tangible contributions.
The public certificates and contributions provide \textbf{peer recognition}.
Fourth, we also found that the focus on the mechanics of a
workflow and the ability to interact with the authors can make
reproductions very \textbf{educational}. It also is a different role and, as such, could
be a welcome option for researchers to give back their time to the community.

@@

% question mark indicates we are waiting for person to respond.
We are grateful to the following individuals for discussions regarding
the work presented here: Andy Collings, Melissa Harrison, Giuliano
Maciocci, Naomi Penfold, Emmy Tsang (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and
Yuhao (Sebastian) Wang developed code and example certificates.
We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
CODECHECKs, Carlos Granell and Frank Ostermann for contributing
certificates as reproducibility reviewers at the AGILE conference, and 
all authors of auditable workflows for their participation.
We thank Celeste R. Brennecka from the Scientific Editing Service, University of M\"unster, for her editorial review.

>>>>

\textbf{Who's got time for more peer review?} Agree.
The time one spends performing CODECHECKs will likely reduce the time an individual has to work as a reviewer.
However, a CODECHECK is different from peer review.
First, the \textbf{technical nature} of a CODECHECK can provide a more 
clearly defined scope and expectations and thereby time budget compared to conventional 
peer review. 
For example, authors are told what to provide and the codechecker can be told when to stop.
The option for the codechecker to ask for clarification or more 
material can greatly reduce the digging required and thereby
increase \textbf{efficiency}.
Second, the specific skill set of a codechecker allows for \textbf{different groups} to take part in the review process.
ECRs might be attracted to learn more about recent methods, peer review, and reproducibility practices.
Research software experts (Research Software Engineers, RSE\footnote{\url{https://en.wikipedia.org/wiki/Research_software_engineering}}) who might
not regularly be involved in writing or reviewing papers might be interested in increasing their connection with scholarly practices.
An extra codechecker may simplify the matchmaking an editor has to do
when identifying suitable reviewers for a submission, as technical and
topical expertise can be provided by different people who do not even have to be traditional peer reviewers.
In fact, the codechecker could even be a software professional employed by a publisher.
Third, recall that we expect CODECHECKs not to be anonymous and always
publicly deposited, which is common practice for reviews of few journals only.
With code and workflows, there also is a good chance that a codechecker's
feedback can directly impact and improve the author's 
work, which may make for a rewarding experience or even provide tangible contributions.
The public certificates and contributions provide \textbf{peer recognition}.
Fourth, we also found that the focus on the mechanics of a
workflow and the ability to interact with the authors can make
reproductions very \textbf{educational}. It also is a different role and, as such, could
be a welcome option for researchers to give back their time to the community.

@@

We are grateful to the following individuals for discussions regarding
the work presented here: Andy Collings, Melissa Harrison, Giuliano
Maciocci, Naomi Penfold, Emmy Tsang (eLife), Rebecca Kirk (PLOS Computational Biology), Scott
Edmunds (GigaScience), and Andrew Hufton (Scientific Data). Iain Davies and
Yuhao (Sebastian) Wang developed code and example certificates.
We thank Antonio P\'{a}ez (Journal of Geographical Systems) for enabling
CODECHECKs, Carlos Granell and Frank Ostermann for contributing
certificates as reproducibility reviewers at the AGILE conference, and 
all authors of auditable workflows for their participation.
We thank Celeste R. Brennecka from the Scientific Editing Service, University of M\"unster, for her editorial review.
